---
title: "Converting between graph classes"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Converting-between-graph-classes}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---



```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = FALSE,
  comment = "#>",
  eval = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE
)

```

```{r gt_tables, include = FALSE, eval = TRUE, file='create_tables.R'}
```

```{=html}
<style type="text/css">
.main-container {
  max-width: 1200px;
  margin-left: auto;
  margin-right: auto;
}
</style>
```
```{=html}
<style type="text/css">
pre {
    border-style: hidden;
}
</style>
```



# Converting between graph classes

As mentioned [above](#snafun-intro), a common network analysis workflow
includes the conversion of graph objects between various formats. The
`igraph` and `statnet` packages provide some very basic functions for
this, that are specific to the types of objects that are used in these
packages themselves:

```{r table_convert, echo = FALSE, eval = TRUE}
table_convert
```

Conversion between various formats is a lot easier with the help of the
`snafun` package. You only need to know some very easy function names to
convert between the common graph classes. Here is an overview of which
function to use for which conversion:

```{r table_to, echo = FALSE, eval = TRUE}
table_to
```
