---
title: "Creating a graph object"
output: rmarkdown::html_document
vignette: >
  %\VignetteIndexEntry{Creating a graph object}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = FALSE,
  comment = "#>",
  eval = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE
)

```

```{r gt_tables, include = FALSE, eval = TRUE, file='create_tables.R'}
```

```{=html}
<style type="text/css">
.main-container {
  max-width: 1200px;
  margin-left: auto;
  margin-right: auto;
}
</style>
```
```{=html}
<style type="text/css">
pre {
    border-style: hidden;
}
</style>
```



# Creating a graph object {#generate}

This is how you create graph objects of class `igraph` or `network`.

```{r table_create, echo = FALSE, eval = TRUE}
table_create
```

<br>

## Additional graph creation functions

The `snafun` package offers a few functions that assist with the
manipulation of graph data in `R`.

-   `snafun::make_edgelist(names = NULL, attribute = NULL)`

The input is a data.frame (`names`) with edge information. The
`attribute` is a vector that contains a node attribute for those
vertices.

The function returns a vector or data.frame that can be read into
`igraph` or `network`.

-   `snafun::make_nodelist(names = NULL, attribute = NULL)`

The input is a data.frame (`names`) with edge information. The
`attribute` is another data.frame that contains the values of those
edges.

The function returns an edgelist that can be read into `igraph` or
`network`.

## Importing UCINET files

If your data come from UCINET, you can import them directly with
`snafun::read_ucinet()`. The function reads common UCINET `DL` text
files and recent `##h`/`##d` UCINET system files and returns an
`igraph` object by default.

```{r ucinet_example}
tmp <- tempfile(fileext = ".dl")

writeLines(
  c(
    "dl n = 4 format = fullmatrix",
    "labels:",
    "Alice, Bob, Carol, Dave",
    "data:",
    "0 1 0 0",
    "1 0 1 0",
    "0 1 0 0",
    "0 0 0 0"
  ),
  con = tmp
)

g_ucinet <- snafun::read_ucinet(tmp)
g_ucinet
snafun::count_vertices(g_ucinet)
snafun::has_vertexnames(g_ucinet)

unlink(tmp)
```

If you want a different output format, set `graph = "network"`,
`graph = "matrix"`, or `graph = "edgelist"`.




