---
title: "Creating a graph object"
output: rmarkdown::html_document
vignette: >
  %\VignetteIndexEntry{Creating-a-graph-object}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = FALSE,
  comment = "#>",
  eval = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE
)

```

```{r gt_tables, include = FALSE, eval = TRUE, file='create_tables.R'}
```

```{=html}
<style type="text/css">
.main-container {
  max-width: 1200px;
  margin-left: auto;
  margin-right: auto;
}
</style>
```
```{=html}
<style type="text/css">
pre {
    border-style: hidden;
}
</style>
```



# Creating a graph object {#generate}

This is how you create graph objects of class `igraph` or `network`.

```{r table_create, echo = FALSE, eval = TRUE}
table_create
```

<br>

## Additional graph creation functions

The `snafun` package offers a few functions that assist with the
manipulation of graph data in `R`.

-   `snafun::make_edgelist(names = NULL, attribute = NULL)`

The input is a data.frame (`names`) with edge information. The
`attribute` is a vector that contains a node attribute for those
vertices.

The function returns a vector or data.frame that can be read into
`igraph` or `network`.

-   `snafun::make_nodelist(names = NULL, attribute = NULL)`

The input is a data.frame (`names`) with edge information. The
`attribute` is another data.frame that contains the values of those
edges.

The function returns an edgelist that can be read into `igraph` or
`network`.




