---
title: "Dyad-level indices"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Dyad-level indices}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = FALSE,
  comment = "#>",
  eval = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE
)

```

```{r gt_tables, include = FALSE, eval = TRUE, file='create_tables.R'}
```

```{=html}
<style type="text/css">
.main-container {
  max-width: 1800px;
  margin-left: auto;
  margin-right: auto;
}
</style>
```
```{=html}
<style type="text/css">
pre {
    border-style: hidden;
}
</style>
```



# Dyad-level indices

Here are the functions for several of the dyad-level indices you will
want to use in this course.

```{r table_dyad, echo = FALSE, eval = TRUE}
table_dyads
```

## Equivalence as a dyadic idea

Equivalence measures compare vertices in pairs. So even though they are
often used to say something about *roles* or *positions*, the output is
still dyadic: for each pair of vertices, you get a score saying how
similar they are.

In `snafun`, the two main options are:

- `snafun::d_structural_equivalence()`: do two vertices have the same
  ties to the same alters?
- `snafun::d_regular_equivalence()`: do two vertices occupy the same
  kind of position in the network, even if they are not tied to exactly
  the same alters?

Here is a simple comparison.

```{r equivalence_dyad_example, eval = TRUE}
g <- snafun::create_manual_graph(
  A -+ B,
  A -+ C,
  D -+ E,
  D -+ F,
  B -+ G,
  C -+ G,
  E -+ H,
  F -+ H
)

s_eq <- snafun::d_structural_equivalence(g)
r_eq <- snafun::d_regular_equivalence(g)

s_eq
r_eq
```

In this graph:

- `B` and `C` are structurally equivalent, because they have the same
  pattern of ties.
- `E` and `F` are structurally equivalent for the same reason.
- `A` and `D` are not structurally equivalent, because they are tied to
  different alters.
- But `A` and `D` are regularly equivalent, because they occupy the
  same kind of source-like role in the graph.

That is the main teaching distinction:

- structural equivalence is about *same neighbors*;
- regular equivalence is about *same role*.
