---
title: "Graph level indices"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Graph level indices}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = FALSE,
  comment = "#>",
  eval = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE
)

```

```{r gt_tables, include = FALSE, eval = TRUE, file='../inst/extdata/create_tables.R'}
```

```{=html}
<style type="text/css">
.main-container {
  max-width: 1800px;
  margin-left: auto;
  margin-right: auto;
}
</style>
```
```{=html}
<style type="text/css">
pre {
    border-style: hidden;
}
</style>
```



# Graph level indices

Below you will find a table to determine the foundational indices of a
graph at the graph-level.

```{r table_graph, echo = FALSE, eval = TRUE}
table_graph
```

## Communities and other subgroups

It is informative to know that the `snafun` functions to extract
communities yield results that can be scrutinized by `igraph`. The
`snafun` package is smart enough to do this, regardless of whether the
original input graph was of class `igraph` or `network`. Really handy.

Here's an example.

```{r walktrap, echo = TRUE, eval = TRUE}
# generate a random directed graph with 20 vertices and 30 edges
g <- snafun::create_random_graph(20, "gnm", m = 30)

# determine the walktrap communities
walk <- snafun::extract_comm_walktrap(g)
print(walk)

# get the modularity score
snafun::extract_comm_modularity(walk)

# membership of each vertex
snafun::extract_comm_membership(walk)

# number of communities
snafun::count_communities(walk)

# size of each community
snafun::extract_comm_sizes(walk)

# which edge connects multiple communities
snafun::extract_comm_crossing(walk, g)

# plot the network, highlighting the communities
snafun::plot_communities(walk, g)
```

If you are so inclined, you can plot the community division as a
dendrogram, as follows:

```{r dendro, echo = TRUE, eval = TRUE}
snafun::plot_comm_dendrogram(walk)
```

## Writing community results back to the graph

Community detection is often only the first step. In practice, you
usually want to attach the results back to the network so you can use
them in plots, tables, and follow-up analyses.

`snafun` makes this easy:

- `snafun::add_comm_membership()` writes the community membership of
  each vertex back as a vertex attribute.
- `snafun::add_comm_crossing()` writes whether an edge runs between two
  different communities back as an edge attribute.

```{r community_workflow, echo = TRUE, eval = TRUE}
g_comm <- snafun::add_comm_membership(g, walk, attr = "community")
g_comm <- snafun::add_comm_crossing(g_comm, walk, attr = "crossing")

snafun::extract_vertex_attribute(g_comm, "community")
snafun::extract_edge_attribute(g_comm, "crossing")
```

This workflow is especially useful when you want to:

- color vertices by community in a later plot;
- identify bridging edges between communities; or
- compare community structure to other vertex attributes.

## Evaluating community quality

After finding a community structure, you may also want to evaluate how
well-separated the detected communities are. `snafun` can do this with
`evaluate_communities()`.

The function returns a community-level table with:

- `Silhouette`: the average silhouette width of the vertices in that
  community; higher values indicate cleaner separation.
- `Modularity`: the modularity of the full partition; higher values
  indicate relatively many ties within communities and relatively few
  ties between communities.
- `Conductance`: how strongly a community is tied to the rest of the
  graph; lower values are usually better.
- `Density`: how densely connected the vertices are within the
  community itself.

```{r evaluate_communities, echo = TRUE, eval = TRUE}
evls <- snafun::evaluate_communities(walk, g)
evls
```

If you also want the silhouette width for each individual vertex, set
`sil_width = TRUE`.

```{r evaluate_communities_sil, echo = TRUE, eval = TRUE}
evls2 <- snafun::evaluate_communities(walk, g, sil_width = TRUE)
head(evls2$sil_width)
```

You can also plot the silhouette widths directly.

```{r evaluate_communities_plot, echo = TRUE, eval = TRUE, fig.width = 10, fig.height = 7}
plot(evls)
```

In that plot:

- bars above `0` indicate vertices that fit their own community better
  than neighbouring communities;
- bars near `0` indicate ambiguous or boundary cases;
- bars below `0` indicate vertices that may fit another community
  better than the one they were assigned to;
- the red line shows the overall average silhouette width.

# Vertex-level indices

Here are the functions to determine many of the vertex-level indices you
will want to use in this course.

```{r table_vertex, echo = FALSE, eval = TRUE}
table_vertices
```
