---
title: "Graph level indices"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Graph-level-indices}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = FALSE,
  comment = "#>",
  eval = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE
)

```

```{r gt_tables, include = FALSE, eval = TRUE, file='create_tables.R'}
```

```{=html}
<style type="text/css">
.main-container {
  max-width: 1800px;
  margin-left: auto;
  margin-right: auto;
}
</style>
```
```{=html}
<style type="text/css">
pre {
    border-style: hidden;
}
</style>
```



# Graph level indices

Below you will find a table to determine the foundational indices of a
graph at the graph-level.

```{r table_graph, echo = FALSE, eval = TRUE}
table_graph
```

## Communities and other subgroups

It is informative to know that the `snafun` functions to extract
communities yield results that can be scrutinized by `igraph`. The
`snafun` package is smart enough to do this, regardless of whether the
original input graph was of class `igraph` or `network`. Really handy.

Here's an example.

```{r walktrap, echo = TRUE, eval = TRUE}
# generate a random directed graph with 20 vertices and 30 edges
g <- snafun::create_random_graph(20, "gnm", m = 30)

# determine the walktrap communities
walk <- snafun::extract_comm_walktrap(g)
print(walk)

# get the modularity score
igraph::modularity(walk)

# who is member of which community
igraph::communities(walk)

# which community is a vertex member of
igraph::membership(walk)

# number of communities
length(walk)

# size of each community
igraph::sizes(walk)

# which edge connects multiple communities
igraph::crossing(walk, g)

# plot the network, highlighting the communities
plot(walk, g)
```

If you are so inclined, you can plot the community division as a
dendrogram, as follows:

```{r dendro, echo = TRUE, eval = TRUE}
snafun::plot_comm_dendrogram(walk)
```

# Vertex-level indices

Here are the functions to determine many of the vertex-level indices you
will want to use in this course.

```{r table_vertex, echo = FALSE, eval = TRUE}
table_vertices
```
