---
title: "Introduction"
output: rmarkdown::html_document
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = FALSE,
  comment = "#>",
  eval = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE
)


```



<br><br>

# Main packages in the SNA4DS course {#main_packages}

## Overview of `igraph` and `network` graph objects

There are two main packages for basic graph generation and manipulation:
the `igraph` package and the `statnet` package. Actually, `statnet` is a
suite of packages that work together. In this course, we will will make
use of several packages from the `statnet` suite.

The `igraph` package creates a graph object of type `igraph`. The
`statnet` suite creates a graph object of type `network`. There are many
things you can do in both packages. Both packages can generate graphs
and do basic manipulation, so here you should just use the package whose
API you like best. The `igraph` package provides more mathematical
functions to apply to the graph data and the `statnet` suite provides
loads of statistical models that the `igraph` package does not do.

## The `snafun` package {#snafun-intro}

The `igraph` package and `statnet` suite are jointly very powerful and
can support much of your analyses of network data. However, as you read
above, they each require graph objects that have specific structures and
they can't deal with a graph object that has a different structure. So,
if you want to use the functions from both the `igraph` and the `sna`
packages, you need network data in `igraph` format (for the `igraph`
package), in `network` format (for the `network` package and some of the
`sna` package) and in `matrix` format (for many of the functions in the
`sna` package). In other words, you will have to convert your data
between these formats and you also have to deal with the differing API's
between these various packages.

Believe it or not, this is a pain and quite annoying.

**THE `snafun` PACKAGE TO THE RESCUE!**

The `snafun` package does three things:

-   First, it provides an (fairly) consistent API, so you don't have to
    constantly figure out what a specific argument means for each
    function;
-   Second, most of the functions in the `snafun` package work on both
    objects of class `igraph` or `network`. As a result, you can do what
    you want to do, without bothering with whether the object you work
    on is of class `igraph` or `network`.
-   Third, by removing the pain coming from the constant switching
    between the two groups of packages and their inconsistent API, you
    can now actually focus on the **fun** of network analysis, rather
    than the frustration.

Oh, and there is a *fourth advantage* too: the authors of the `snafun`
package are cool people. So, if you have the need for a new function in
the package, just get in touch with us and we'll see what we can do for
you. 

:-)


# Structural versus regular equivalence {#equivalence}

Two vertices can be similar in more than one way. `snafun` contains
functions for both *structural equivalence* and *regular equivalence*.

- Structural equivalence asks whether two vertices are tied to the same
  alters in the same way.
- Regular equivalence asks whether two vertices occupy the same kind of
  position, even if they are not tied to exactly the same alters.

For students, a useful rule of thumb is this:

- use `snafun::d_structural_equivalence()` when you care about matching
  neighborhoods;
- use `snafun::d_regular_equivalence()` when you care about matching
  roles or positions in the graph.

The regular-equivalence function in `snafun` is based on the iterative
logic of `sna::redist()`, but it returns an *equivalence* matrix between
0 and 1 rather than a distance matrix.

```{r equivalence_examples}
g <- snafun::create_manual_graph(
  A -+ B,
  A -+ C,
  D -+ E,
  D -+ F,
  B -+ G,
  C -+ G,
  E -+ H,
  F -+ H
)

# Same exact ties -> structural equivalence
snafun::d_structural_equivalence(g)

# Same role pattern -> regular equivalence
snafun::d_regular_equivalence(g)

# If weights matter, let them contribute explicitly
g_w <- snafun::add_edge_attributes(
  g,
  attr_name = "weight",
  value = c(2, 1, 2, 1, 3, 3, 3, 3)
)
snafun::d_regular_equivalence(g_w, weights = "weight")
```

## How to read an equivalence matrix

An equivalence matrix is easiest to read as a table of pairwise
similarities.

- Values close to `1` indicate that two vertices are very similar under
  the chosen equivalence concept.
- Values close to `0` indicate that two vertices are very different.
- The diagonal is always `1`, because every vertex is perfectly
  equivalent to itself.

In the example above, the structural-equivalence matrix mainly picks up
vertices with the *same exact outgoing pattern*, such as `B` and `C`,
or `E` and `F`. The regular-equivalence matrix is more forgiving: it
should also reveal that `A` and `D` occupy the same *role* in the
network, even though they are tied to different alters.

So for this graph, you would typically expect the regular-equivalence
matrix to suggest role sets like:

- `A` and `D`: source-like vertices
- `B`, `C`, `E`, and `F`: middle-layer vertices
- `G` and `H`: sink-like vertices

This is one of the main reasons to look at regular equivalence in
teaching settings. It helps students distinguish between:

- *same neighbors*; and
- *same position in the wider graph structure*.

If you want edge strengths to affect those role similarities, use the
`weights` argument explicitly. If you do not specify weights, the
regular-equivalence calculation treats the graph as unweighted.

