---
title: "Manipulating the graph object"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Manipulating the graph object}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = FALSE,
  comment = "#>",
  eval = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE
)

```

```{r gt_tables, include = FALSE, eval = TRUE, file='create_tables.R'}
```

```{=html}
<style type="text/css">
.main-container {
  max-width: 1200px;
  margin-left: auto;
  margin-right: auto;
}
</style>
```
```{=html}
<style type="text/css">
pre {
    border-style: hidden;
}
</style>
```



# Manipulating the graph object {#manipulate}

Once you have a network object, you will want to explore it and access
parts of it or extract info from it.

Here is a table with some of the functions you'll use all the time in
any SNA project.

```{r table_manipulate, echo = FALSE, eval = TRUE}
table_manipulate
```

One common manipulation for two-mode data is to create one-mode
projections. `snafun::make_projection()` now offers the same core
behaviour as `igraph::bipartite_projection()`, but with a consistent API
across `igraph`, `network`, edge lists, and incidence matrices.

The clearest way to request a projection is with `which = "row"` or
`which = "column"`, referring to the row and column partitions of the
underlying incidence matrix. If you want both projections, use
`which = "both"`.

```{r projection_example, eval = TRUE}
incidence <- matrix(
  c(
    1, 1,
    1, 1,
    0, 1
  ),
  nrow = 3,
  byrow = TRUE,
  dimnames = list(
    c("Alice", "Bob", "Cecil"),
    c("Project1", "Project2")
  )
)

row_projection <- snafun::make_projection(
  incidence,
  which = "row"
)
column_projection <- snafun::make_projection(
  incidence,
  which = "column",
  multiplicity = FALSE
)

row_projection
column_projection
```

The row projection above is weighted by default, because the edge values
count how many events two actors share. Setting `multiplicity = FALSE`
returns a binary projection instead.

