---
title: "Manipulating the graph object"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Manipulating the graph object}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = FALSE,
  comment = "#>",
  eval = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE
)

```

```{r gt_tables, include = FALSE, eval = TRUE, file='../inst/extdata/create_tables.R'}
```

```{=html}
<style type="text/css">
.main-container {
  max-width: 1200px;
  margin-left: auto;
  margin-right: auto;
}
</style>
```
```{=html}
<style type="text/css">
pre {
    border-style: hidden;
}
</style>
```



# Manipulating the graph object {#manipulate}

Once you have a network object, you will want to explore it and access
parts of it or extract info from it.

Here is a table with some of the functions you'll use all the time in
any SNA project.

```{r table_manipulate, echo = FALSE, eval = TRUE}
table_manipulate
```

One common manipulation for two-mode data is to create one-mode
projections. `snafun::make_projection()` now offers the same core
behaviour as `igraph::bipartite_projection()`, but with a consistent API
across `igraph`, `network`, edge lists, and incidence matrices.

The clearest way to request a projection is with `which = "row"` or
`which = "column"`, referring to the row and column partitions of the
underlying incidence matrix. If you want both projections, use
`which = "both"`.

```{r projection_example, eval = TRUE}
incidence <- matrix(
  c(
    1, 1,
    1, 1,
    0, 1
  ),
  nrow = 3,
  byrow = TRUE,
  dimnames = list(
    c("Alice", "Bob", "Cecil"),
    c("Project1", "Project2")
  )
)

row_projection <- snafun::make_projection(
  incidence,
  which = "row"
)
column_projection <- snafun::make_projection(
  incidence,
  which = "column",
  multiplicity = FALSE
)

row_projection
column_projection
```

The row projection above is weighted by default, because the edge values
count how many events two actors share. Setting `multiplicity = FALSE`
returns a binary projection instead.

## Working with multiple edges

Some network data contain multiple edges between the same pair of
vertices. In teaching settings, it is often useful to first inspect
whether those multiple edges are present, and then decide whether they
should be kept or collapsed.

`snafun` provides three simple helpers for this:

- `snafun::has_multiple_edges()` checks whether the graph contains
  parallel edges.
- `snafun::extract_multiple_edges()` returns the ids of the extra
  parallel edges.
- `snafun::remove_multiple_edges()` collapses them to a single edge per
  dyad, while combining edge attributes using explicit rules.

```{r multiple_edges_example, eval = TRUE}
g_multi <- snafun::create_manual_graph(
  A -+ B,
  A -+ B,
  B -+ C,
  simplify = FALSE
)

g_multi <- snafun::add_edge_attributes(
  g_multi,
  attr_name = "weight",
  value = c(2, 5, 1)
)
g_multi <- snafun::add_edge_attributes(
  g_multi,
  attr_name = "type",
  value = c("email", "chat", "email")
)

snafun::has_multiple_edges(g_multi)
snafun::extract_multiple_edges(g_multi)

g_simple <- snafun::remove_multiple_edges(
  g_multi,
  edge_attr_combine = list(weight = "sum", type = "concat_unique")
)

g_simple
snafun::extract_edge_attribute(g_simple, "weight")
snafun::extract_edge_attribute(g_simple, "type")
```

This is especially useful when weights represent repeated interaction.
In that case, collapsing parallel edges with `weight = "sum"` often
gives the most interpretable simple graph.

