---
title: "Statistical Models"
output: rmarkdown::html_document
vignette: >
  %\VignetteIndexEntry{Statistical-models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = FALSE,
  comment = "#>",
  eval = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE
)

```

```{r gt_tables, include = FALSE, eval = TRUE, file='create_tables.R'}
```

```{=html}
<style type="text/css">
.main-container {
  max-width: 1200px;
  margin-left: auto;
  margin-right: auto;
}
</style>
```
```{=html}
<style type="text/css">
pre {
    border-style: hidden;
}
</style>
```





# Statistical models

## Overview table

Here is an overview of the statistical models discussed in the course.

<br>

```{r, echo = FALSE, eval = TRUE}
table_models
```

<br>

## Network autocorrelation models

The network autocorrelation model is run through the `snafun::stat_nam`
function. The implementation is appropriate for continuous dependent
variables.

The basic function call is as follows:

```{r}
snafun::stat_nam(
  formula,
  data = list(),
  W,
  W2 = NULL,
  model = c("lag", "error", "combined"),
  na.action,
  Durbin = FALSE,
  quiet = TRUE,
  zero.policy = TRUE,
  check_vars = TRUE
)
```

Here,

-   `formula` is a formula of the form `DEP ~ EXO1 + EXO2 + EXO3` or
    `DEP ~ .` This follows the standard way of formulating formulas
    in R. Run `?stats::formula` in your console for details.

-   `data` is a `data.frame` or a `list` containing the variables to be
    included into the model.

-   `W` is a matrix of the same dimension as the network, containing the
    weights that drive the network influence process (for the `lag`
    model) or the network disturbances process (for the `error` model).

-   `W2` is a matrix of the same dimension as the network, containing
    the weights that drive the network disturbances when running a
    `combined` model.

An intercept is added to the model by default, it is best to NOT include
this as a separate exogenous variable.

The weigh matrix (matrices) is (are) row-normalized by this function. If
you want to run the model using non-standardized weight matrices, use
the `sna::lnam` function instead.

The `snafun` package provides a useful `stat_nam_summary` method (that
shows you an overview of the results) and a `plot_nam` method (that you
use to check model assumptions).

<br>

## Conditional Uniform graphs (CUG)

There are two methods to perform a conditional Uniform graph test.

The first is to generate the graphs manually and calculate the measures
on each graph. Generation of these graphs can be done using
`snafun::create_random_graph` (which conditions on size and density).
The equivalent functions in `sna` are `sna::rgraph` and `sna::rgnm`
(note that these generate a matrix, rather than an `igraph` or `network`
graph). See [the data generation table](#generate) for these functions.

The second approach is to use a function that does the graph generation
and computes the network measure for you. The preferred is
`sna::cugtest`, which is specified as follows:

```{r}
sna::cug.test(g, FUN, mode = c("digraph", "graph"), cmode = c("size",
    "edges", "dyad.census"), reps = 1000,
    ignore.eval = TRUE, FUN.args = list())
```

See the `sna` help function for details.

Here

-   `FUN` is the function that needs to be calculated on each graph

-   `FUN.args` contains any arguments that are required for the function
    you specified in `FUN`

-   `cmode` determines the type of graphs that are drawn (ie. what you
    condition on). The options are

    -   "size": this generates graphs with a particular size and density
        0.5. You rarely want this.

    -   "edges": this conditions on a specific edge count (or an exact
        edge value distribution)

    -   "dyad.census": this conditions on a dyad census (or dyad value
        distribution)

For example, in order to test whether the transitivity in your graph `g`
is exceptional for a network of the same size and density as in `g`, you
would run

```{r}
sna::cug.test(g, sna::gtrans, cmode = "edges")
```

It is wise to always explicitly tell the function whether your graph is
directed or not, so a better way to specify the previous function is

```{r}
sna::cug.test(g, mode = "graph", FUN = sna::gtrans,
              cmode = "edges", reps = 1000,
              FUN.args = list(mode = "graph"))
```

Testing the betweenness centralization of you network `g` could be
performed as follows, again conditioning on size and density:

```{r}
sna::cug.test(g,
              sna::centralization,
              FUN.arg=list(FUN = sna::betweenness),
              mode="graph",
              cmode="edges")
```

There is also a useful `plot` method for the result of the CUG test.

<br>

The huuuuuge downside of the `cug.test` function is that it only takes
specific input graps (perferably in matrix-format) and only allows one
to test with functions implemented within the `sna` package. However,
`snafun` can fix that. You only need to construct a simple function with
the following structure <small>(yes, always start the function with
`x <- snafun::fix_cug_input(x, directed = directed)`, this fixes some
peculiarities related to the `sna::cug.test` function)</small>:

```{r}
trans_f <- function(x, directed = FALSE) {
  x <- snafun::fix_cug_input(x, directed = directed)
  snafun::g_transitivity(x)
}
```

and you can now test for transitivity with ANY graph that works with
`snafun::g_transitivity(x)` (which includes most formats you'll work
with). Then, perform the CUG test as follows:

```{r}
sna::cug.test(graph_object, 
              mode = "graph", 
              FUN = trans_f, 
              cmode = "edges", 
              reps = 200)
```

Et voila. It simply works.

Wanna use a more involved measure that doesn't exist in the `sna`
package anyway? Sure, this is how you run a CUG test on the number of
walktrap communities in the graph:

```{r}
walk_num <- function(x, directed = FALSE) {
  x <- snafun::fix_cug_input(x, directed = directed)
  snafun::extract_comm_walktrap(x) |> length()
}

sna::cug.test(graph_object, mode = "graph", FUN = walk_num, cmode = "edges", reps = 200)
```

As said, `snafun` puts the *FUN* straight into SNA.

<br>

## QAP test

There are two methods to perform a QAP test.

The first is to manually permute the graph. Generation of these graphs
can be done using `igraph::permute` or `sna::rmperm`. See [the data
generation table](#generate) for these functions.

The second approach is to use a function that does the graph permutation
and computes the required measure (typically a correlation) for you. The
preferred is `sna::qaptest`, which is specified as follows:

```{r}
sna::qaptest(g, FUN, reps = 1000, ...)
```

See the `sna` help function for details.

Here

-   `FUN` is the function that needs to be calculated after each
    permutation

-   `...` contains any arguments that are required for the function you
    specified in `FUN`

Typically, you want to test the correlation between two graphs, as
follows:

```{r}
sna::qaptest(list(firstNetwork, secondNetwork),
             FUN = sna::gcor, reps = 1000,
             g1 = 1, g2 = 2)
```

There is a useful `summary` method and a `plot` method for the output of
the function.

<br>

## QAP linear regression

QAP linear regression is performed through the `sna::netlm` function.
The function looks as follows:

```{r}
sna::netlm(y, x, intercept = TRUE, mode = "digraph",
    nullhyp = "qapspp", reps = 1000)
```

Make sure to always set `intercept = TRUE` and `nullhyp = "qapspp"`. For
small networks, 1000 replications should be enough, for larger networks
you should typically use a higher number (say, 2000).

As an example, this is how you specify a model where graph `g` is
modeled as a linear function of graphs `g1`, `g2`, and `g3`.

```{r}
mod <- sna::netlm(y = g, x = list(g1, g2, g3), intercept = TRUE,
                              nullhyp = 'qapspp', reps = 1001)
mod$names <- c("Intcpt", "Net1", "Net2", "Net3")
summary(mod)
```

It is wise to add the names of the networks to the output object, like
you see above. That is not strictly necessary, but it makes the output
of the function easier to read.

<br>

## QAP logistic regression

QAP logistic regression is performed through the `sna::netlogit`
function. The function looks as follows:

```{r}
sna::netlogit(y, x, intercept = TRUE, mode = "digraph",
    nullhyp = "qapspp", reps = 1000)
```

Make sure to always set `intercept = TRUE` and `nullhyp = "qapspp"`. For
small networks, 1000 replications should be enough, for larger networks
you should typically use a higher number (say, 2000).

As an example, this is how you specify a model where binary graph `g` is
modeled as a function of graphs `g1`, `g2`, and `g3`.

```{r netlogit1, exercise = TRUE}
mod <- sna::netlogit(g, list(g1, g2, g3),
                     intercept = TRUE,
                     nullhyp = "qapspp", reps = 1001)
mod$names <- c("Intcpt", "Net1", "Net2", "Net3")
summary(mod)
```
