---
title: "Statistical Models"
output: rmarkdown::html_document
vignette: >
  %\VignetteIndexEntry{Statistical Models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = FALSE,
  comment = "#>",
  eval = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE
)

```

```{r gt_tables, include = FALSE, eval = TRUE, file='create_tables.R'}
```

```{=html}
<style type="text/css">
.main-container {
  max-width: 1200px;
  margin-left: auto;
  margin-right: auto;
}
</style>
```
```{=html}
<style type="text/css">
pre {
    border-style: hidden;
}
</style>
```





# Statistical models

## Overview table

Here is an overview of the statistical models discussed in the course.

<br>

```{r, echo = FALSE, eval = TRUE}
table_models
```

<br>

## Network autocorrelation models

The network autocorrelation model is run through the `snafun::stat_nam`
function. The implementation is appropriate for continuous dependent
variables.

The basic function call is as follows:

```{r}
snafun::stat_nam(
  formula,
  data = list(),
  W,
  W2 = NULL,
  model = c("lag", "error", "combined"),
  na.action,
  Durbin = FALSE,
  quiet = TRUE,
  zero.policy = TRUE,
  check_vars = TRUE
)
```

Here,

-   `formula` is a formula of the form `DEP ~ EXO1 + EXO2 + EXO3` or
    `DEP ~ .` This follows the standard way of formulating formulas
    in R. Run `?stats::formula` in your console for details.

-   `data` is a `data.frame` or a `list` containing the variables to be
    included into the model.

-   `W` is a matrix of the same dimension as the network, containing the
    weights that drive the network influence process (for the `lag`
    model) or the network disturbances process (for the `error` model).

-   `W2` is a matrix of the same dimension as the network, containing
    the weights that drive the network disturbances when running a
    `combined` model.

An intercept is added to the model by default, it is best to NOT include
this as a separate exogenous variable.

The weigh matrix (matrices) is (are) row-normalized by this function. If
you want to run the model using non-standardized weight matrices, use
the `sna::lnam` function instead.

The preprocessing of the weight matrix deserves a bit of extra attention:

- `W` and `W2` may be supplied as matrices, `igraph` objects, `network`
  objects, or edgelists/data frames, but after conversion they must
  represent square one-mode weight matrices.
- Non-zero rows are row-standardized internally. If a row already sums to
  one, it is left alone; if not, `snafun::stat_nam()` standardizes it for
  you and reports that via a message.
- Zero rows are allowed when `zero.policy = TRUE`. That is the natural way
  to represent isolates or actors without outgoing influence ties.
- Diagonal values are kept. So if you include self-weights, they are part
  of the fitted NAM instead of being dropped silently.
- Edge weights are also kept during conversion. For edgelists, columns 1
  and 2 are interpreted as sender and receiver, and column 3 (if present)
  is used as the numeric weight.

The `snafun` package provides a useful `stat_nam_summary` method (that
shows you an overview of the results) and a `plot_nam` method (that you
use to check model assumptions).

<br>

## Conditional Uniform graphs (CUG)

There are two methods to perform a conditional Uniform graph test.

The first is to generate the graphs manually and calculate the measures
on each graph. Generation of these graphs can be done using
`snafun::create_random_graph` (which conditions on size and density) or
`snafun::create_census_graph` (which conditions on the dyad census). The
equivalent functions in `sna` are `sna::rgraph`, `sna::rgnm`, and
`sna::rguman` (note that these generate matrices rather than `igraph` or
`network` objects). See [the data generation table](#generate) for these
functions.

The second, and usually more convenient, approach is to use
`snafun::stat_cug()`. This function generalizes `sna::cug.test()` to the
graph classes supported in `snafun` and lets you use `sna`, `igraph`,
`snafun`, or custom statistics directly.

```{r}
snafun::stat_cug(
  x,
  FUN,
  mode = c("auto", "digraph", "graph"),
  cmode = c("size", "edges", "dyad.census"),
  diag = FALSE,
  reps = 1000,
  graph = c("auto", "same", "igraph", "network", "matrix", "edgelist"),
  ignore.eval = TRUE,
  FUN.args = list()
)
```

Here

-   `FUN` is the function that needs to be calculated on each graph

-   `FUN.args` contains any additional arguments for the function you
    specified in `FUN`

-   `cmode` determines the type of graphs that are drawn, i.e. what you
    condition on. The options are

    -   `"size"`: this generates graphs with the same size and tie
        probability 0.5. You rarely want this.

    -   `"edges"`: this conditions on a specific edge count

    -   `"dyad.census"`: this conditions on a dyad census

-   `graph` determines in which graph class `FUN` is evaluated. If you
    leave it at `"auto"`, `snafun` first tries the original graph class
    and then tries the other supported classes until `FUN` works.

-   `ignore.eval = TRUE` means that current CUG tests in `snafun` are
    structural and binary: weights are ignored and every non-zero tie is
    treated as present.

For example, in order to test whether the transitivity in your graph `g`
is exceptional for a network of the same size and edge count as in `g`,
you would run:

```{r}
snafun::stat_cug(g, FUN = snafun::g_transitivity, cmode = "edges")
```

If your graph is undirected, it is still wise to say so explicitly:

```{r}
snafun::stat_cug(
  g,
  FUN = snafun::g_transitivity,
  mode = "graph",
  cmode = "edges",
  reps = 1000
)
```

You can also use `igraph` functions directly, even when the observed
network is not stored as an `igraph` object. For example:

```{r}
snafun::stat_cug(
  graph_object,
  FUN = snafun::g_transitivity,
  mode = "graph",
  cmode = "edges",
  reps = 200,
  graph = "same"
)
```

Or a more involved custom measure that does not exist in `sna` at all,
such as the number of walktrap communities:

```{r}
walk_num <- function(x) {
  length(snafun::extract_comm_walktrap(x))
}

snafun::stat_cug(
  graph_object,
  FUN = walk_num,
  mode = "graph",
  cmode = "edges",
  reps = 200,
  graph = "igraph"
)
```

The output can be printed directly and there is also a `plot()` method
for the result.

<br>

## QAP test

There are two methods to perform a QAP test.

The first is to manually permute the graph. Generation of these graphs
can be done using backend package functions such as `igraph::permute` or
`sna::rmperm`. `snafun` does not yet have a dedicated permutation helper
for this, so those backend functions are still useful when you want to
build a custom QAP procedure by hand.

The second approach, and usually the better one in `snafun`, is to use
`snafun::stat_qap_cor()`. This function aligns graph objects by vertex
names, converts `igraph`, `network`, `matrix`, and `data.frame`
edgelists automatically, and then performs a QAP correlation test.

```{r}
snafun::stat_qap_cor(
  x,
  y,
  controls = NULL,
  method = c("pearson", "spearman"),
  reps = 1000,
  diagonal = FALSE,
  directed = c("auto", "directed", "undirected"),
  alternative = c("two.sided", "greater", "less"),
  seed = NULL
)
```

Here

-   `x` and `y` are the two networks whose association you want to test.

-   `controls` optionally contains one or more control graphs, so the
    function computes a partial QAP correlation.

-   `method` chooses between Pearson and Spearman correlation.

-   `directed` and `diagonal` determine which cells of the adjacency
    matrix are included.

Typically, you want to test the correlation between two graphs, as
follows:

```{r}
snafun::stat_qap_cor(firstNetwork, secondNetwork, reps = 1000)
```

If you want to control for a third network, you can do so directly:

```{r}
snafun::stat_qap_cor(
  firstNetwork,
  secondNetwork,
  controls = controlNetwork,
  reps = 1000
)
```

There is a useful `summary` method and a `plot` method for the output of
the function.

When you read the output, the most important pieces are:

- the observed correlation;
- the permutation probabilities `Pr(X>Obs)`, `Pr(X<Obs)`, and
  `Pr(X=Obs)`;
- whether the association remains after any control networks are added.

If the observed correlation is clearly positive and `Pr(X>Obs)` is very
small, that means the observed association is stronger than what you
would usually see under random relabelings of the network. If
`controls` are used, the test concerns the *partial* association between
`x` and `y`, after removing the part explained by the control networks.

<br>

## QAP linear regression

QAP linear regression is performed in `snafun` through
`snafun::stat_qap_lm()`. Internally, this function follows the logic of
`sna::netlm()`, but it accepts all common graph representations used in
this course and keeps the predictor names in the output whenever
possible.

```{r}
snafun::stat_qap_lm(
  y,
  x,
  intercept = TRUE,
  diagonal = FALSE,
  directed = c("auto", "directed", "undirected"),
  nullhyp = c("qapspp", "qapy"),
  reps = 1000,
  test.statistic = c("t-value", "beta"),
  seed = NULL
)
```

Make sure to always set `intercept = TRUE` and `nullhyp = "qapspp"`. For
small networks, 1000 replications should be enough, for larger networks
you should typically use a higher number (say, 2000).

As an example, this is how you specify a model where graph `g` is
modeled as a linear function of graphs `g1`, `g2`, and `g3`.

```{r}
mod <- snafun::stat_qap_lm(
  y = g,
  x = list(g1, g2, g3),
  intercept = TRUE,
  nullhyp = "qapspp",
  reps = 1001
)
summary(mod)
```

If you supply a named list or a literal list such as `list(g1, g2, g3)`,
`snafun::stat_qap_lm()` keeps those names in the output. That makes the
results easier to read than the default `x1`, `x2`, `x3` style output.

When reading the regression output, focus on these parts:

- `beta`: the estimated regression coefficient;
- `t.value`: the coefficient relative to its standard error;
- the permutation probabilities for the chosen `test.statistic`.

If you use `test.statistic = "beta"`, you are asking whether the raw
coefficient is unusual under the QAP null model. If you use
`test.statistic = "t-value"`, you are asking whether the coefficient is
unusual *relative to its uncertainty*. For teaching, `beta` is often
the easier quantity to interpret as an effect size, while `t-value`
connects more directly to standard regression logic.

<br>

## QAP logistic regression

QAP logistic regression is performed through
`snafun::stat_qap_logit()`. This is the logistic counterpart of
`snafun::stat_qap_lm()` and accepts the same graph classes.

```{r}
snafun::stat_qap_logit(
  y,
  x,
  intercept = TRUE,
  diagonal = FALSE,
  directed = c("auto", "directed", "undirected"),
  nullhyp = c("qapspp", "qapy"),
  reps = 1000,
  test.statistic = c("z-value", "beta"),
  seed = NULL
)
```

Make sure to always set `intercept = TRUE` and `nullhyp = "qapspp"`. For
small networks, 1000 replications should be enough, for larger networks
you should typically use a higher number (say, 2000).

As an example, this is how you specify a model where binary graph `g` is
modeled as a function of graphs `g1`, `g2`, and `g3`.

```{r netlogit1, exercise = TRUE}
mod <- snafun::stat_qap_logit(
  y = g,
  x = list(g1, g2, g3),
  intercept = TRUE,
  nullhyp = "qapspp",
  reps = 1001
)
summary(mod)
```

The logistic output adds a few extra pieces that are especially useful:

- `exp.beta`: the multiplicative change in the odds for a one-unit
  increase in the predictor;
- the goodness-of-fit block, including null and residual deviance;
- the classification block with `accuracy`, `sensitivity`,
  `specificity`, and `precision`.

As a quick reminder:

- `accuracy` is the overall fraction of correctly classified dyads;
- `sensitivity` is the fraction of true ties that the model predicts as
  ties;
- `specificity` is the fraction of true non-ties that the model predicts
  as non-ties;
- `precision` is the fraction of predicted ties that are actually ties.
