---
title: "Statistical Models"
output: rmarkdown::html_document
vignette: >
  %\VignetteIndexEntry{Statistical Models}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  echo = FALSE,
  comment = "#>",
  eval = FALSE,
  echo = TRUE,
  message = FALSE, 
  warning = FALSE
)

```

```{r gt_tables, include = FALSE, eval = TRUE, file='create_tables.R'}
```

```{=html}
<style type="text/css">
.main-container {
  max-width: 1200px;
  margin-left: auto;
  margin-right: auto;
}
</style>
```
```{=html}
<style type="text/css">
pre {
    border-style: hidden;
}
</style>
```





# Statistical models

## Overview table

Here is an overview of the statistical models discussed in the course.

<br>

```{r, echo = FALSE, eval = TRUE}
table_models
```

<br>

## Network autocorrelation models

The network autocorrelation model is run through the `snafun::stat_nam`
function. The implementation is appropriate for continuous dependent
variables.

The basic function call is as follows:

```{r}
snafun::stat_nam(
  formula,
  data = list(),
  W,
  W2 = NULL,
  model = c("lag", "error", "combined"),
  na.action,
  Durbin = FALSE,
  quiet = TRUE,
  zero.policy = TRUE,
  check_vars = TRUE
)
```

Here,

-   `formula` is a formula of the form `DEP ~ EXO1 + EXO2 + EXO3` or
    `DEP ~ .` This follows the standard way of formulating formulas
    in R. Run `?stats::formula` in your console for details.

-   `data` is a `data.frame` or a `list` containing the variables to be
    included into the model.

-   `W` is a matrix of the same dimension as the network, containing the
    weights that drive the network influence process (for the `lag`
    model) or the network disturbances process (for the `error` model).

-   `W2` is a matrix of the same dimension as the network, containing
    the weights that drive the network disturbances when running a
    `combined` model.

An intercept is added to the model by default, it is best to NOT include
this as a separate exogenous variable.

The weigh matrix (matrices) is (are) row-normalized by this function. If
you want to run the model using non-standardized weight matrices, use
the `sna::lnam` function instead.

The preprocessing of the weight matrix deserves a bit of extra attention:

- `W` and `W2` may be supplied as matrices, `igraph` objects, `network`
  objects, or edgelists/data frames, but after conversion they must
  represent square one-mode weight matrices.
- Non-zero rows are row-standardized internally. If a row already sums to
  one, it is left alone; if not, `snafun::stat_nam()` standardizes it for
  you and reports that via a message.
- Zero rows are allowed when `zero.policy = TRUE`. That is the natural way
  to represent isolates or actors without outgoing influence ties.
- Diagonal values are kept. So if you include self-weights, they are part
  of the fitted NAM instead of being dropped silently.
- Edge weights are also kept during conversion. For edgelists, columns 1
  and 2 are interpreted as sender and receiver, and column 3 (if present)
  is used as the numeric weight.

The `snafun` package provides a useful `stat_nam_summary` method (that
shows you an overview of the results) and a `plot_nam` method (that you
use to check model assumptions).

<br>

## Conditional Uniform graphs (CUG)

There are two methods to perform a conditional Uniform graph test.

The first is to generate the graphs manually and calculate the measures
on each graph. Generation of these graphs can be done using
`snafun::create_random_graph` (which conditions on size and density) or
`snafun::create_census_graph` (which conditions on the dyad census). The
equivalent functions in `sna` are `sna::rgraph`, `sna::rgnm`, and
`sna::rguman` (note that these generate matrices rather than `igraph` or
`network` objects). See [the data generation table](#generate) for these
functions.

The second, and usually more convenient, approach is to use
`snafun::stat_cug()`. This function generalizes `sna::cug.test()` to the
graph classes supported in `snafun` and lets you use `sna`, `igraph`,
`snafun`, or custom statistics directly.

```{r}
snafun::stat_cug(
  x,
  FUN,
  mode = c("auto", "digraph", "graph"),
  cmode = c("size", "edges", "dyad.census"),
  diag = FALSE,
  reps = 1000,
  graph = c("auto", "same", "igraph", "network", "matrix", "edgelist"),
  ignore.eval = TRUE,
  FUN.args = list()
)
```

Here

-   `FUN` is the function that needs to be calculated on each graph

-   `FUN.args` contains any additional arguments for the function you
    specified in `FUN`

-   `cmode` determines the type of graphs that are drawn, i.e. what you
    condition on. The options are

    -   `"size"`: this generates graphs with the same size and tie
        probability 0.5. You rarely want this.

    -   `"edges"`: this conditions on a specific edge count

    -   `"dyad.census"`: this conditions on a dyad census

-   `graph` determines in which graph class `FUN` is evaluated. If you
    leave it at `"auto"`, `snafun` first tries the original graph class
    and then tries the other supported classes until `FUN` works.

-   `ignore.eval = TRUE` means that current CUG tests in `snafun` are
    structural and binary: weights are ignored and every non-zero tie is
    treated as present.

For example, in order to test whether the transitivity in your graph `g`
is exceptional for a network of the same size and edge count as in `g`,
you would run:

```{r}
snafun::stat_cug(g, FUN = snafun::g_transitivity, cmode = "edges")
```

If your graph is undirected, it is still wise to say so explicitly:

```{r}
snafun::stat_cug(
  g,
  FUN = snafun::g_transitivity,
  mode = "graph",
  cmode = "edges",
  reps = 1000
)
```

You can also use `igraph` functions directly, even when the observed
network is not stored as an `igraph` object. For example:

```{r}
snafun::stat_cug(
  graph_object,
  FUN = igraph::transitivity,
  mode = "graph",
  cmode = "edges",
  reps = 200,
  graph = "igraph"
)
```

Or a more involved custom measure that does not exist in `sna` at all,
such as the number of walktrap communities:

```{r}
walk_num <- function(x) {
  length(snafun::extract_comm_walktrap(x))
}

snafun::stat_cug(
  graph_object,
  FUN = walk_num,
  mode = "graph",
  cmode = "edges",
  reps = 200,
  graph = "igraph"
)
```

The output can be printed directly and there is also a `plot()` method
for the result.

<br>

## QAP test

There are two methods to perform a QAP test.

The first is to manually permute the graph. Generation of these graphs
can be done using `igraph::permute` or `sna::rmperm`. See [the data
generation table](#generate) for these functions.

The second approach is to use a function that does the graph permutation
and computes the required measure (typically a correlation) for you. The
preferred is `sna::qaptest`, which is specified as follows:

```{r}
sna::qaptest(g, FUN, reps = 1000, ...)
```

See the `sna` help function for details.

Here

-   `FUN` is the function that needs to be calculated after each
    permutation

-   `...` contains any arguments that are required for the function you
    specified in `FUN`

Typically, you want to test the correlation between two graphs, as
follows:

```{r}
sna::qaptest(list(firstNetwork, secondNetwork),
             FUN = sna::gcor, reps = 1000,
             g1 = 1, g2 = 2)
```

There is a useful `summary` method and a `plot` method for the output of
the function.

<br>

## QAP linear regression

QAP linear regression is performed through the `sna::netlm` function.
The function looks as follows:

```{r}
sna::netlm(y, x, intercept = TRUE, mode = "digraph",
    nullhyp = "qapspp", reps = 1000)
```

Make sure to always set `intercept = TRUE` and `nullhyp = "qapspp"`. For
small networks, 1000 replications should be enough, for larger networks
you should typically use a higher number (say, 2000).

As an example, this is how you specify a model where graph `g` is
modeled as a linear function of graphs `g1`, `g2`, and `g3`.

```{r}
mod <- sna::netlm(y = g, x = list(g1, g2, g3), intercept = TRUE,
                              nullhyp = 'qapspp', reps = 1001)
mod$names <- c("Intcpt", "Net1", "Net2", "Net3")
summary(mod)
```

It is wise to add the names of the networks to the output object, like
you see above. That is not strictly necessary, but it makes the output
of the function easier to read.

<br>

## QAP logistic regression

QAP logistic regression is performed through the `sna::netlogit`
function. The function looks as follows:

```{r}
sna::netlogit(y, x, intercept = TRUE, mode = "digraph",
    nullhyp = "qapspp", reps = 1000)
```

Make sure to always set `intercept = TRUE` and `nullhyp = "qapspp"`. For
small networks, 1000 replications should be enough, for larger networks
you should typically use a higher number (say, 2000).

As an example, this is how you specify a model where binary graph `g` is
modeled as a function of graphs `g1`, `g2`, and `g3`.

```{r netlogit1, exercise = TRUE}
mod <- sna::netlogit(g, list(g1, g2, g3),
                     intercept = TRUE,
                     nullhyp = "qapspp", reps = 1001)
mod$names <- c("Intcpt", "Net1", "Net2", "Net3")
summary(mod)
```
